### 2017-04-23 paper/jfp-submission/safe.tex
more text

The remaining 473 programs for which we could not produce a witness did
not admit such an automatic diagnosis, so we selected a random sample of
50 programs and manually searched for a witness.
%
We found only 8 cases (or 16\%) where a witness existed, but \toolname
was unable to discover it.
%
Five involved synthesizing a \emph{pair} of specially-crafted inputs
that would result in the function returning values of incompatible
types.
%
The other three required synthesizing an input that would trigger a
particular path through the program, and would likely have been caught
by symbolic execution.
%
Overall, it appears that our random instantiation of holes is well-suited
to finding type errors in student programs.

We categorized the remaining 42 failures into four groups.

@@

\paragraph{Bad Function Calls}
Fourteen programs contained a function call where the student had
supplied incorrect inputs, yet no error was triggered.
%
Consider, for example the following |assoc| function, which
looks up a key in an \emph{association list} and returns a default if
the key cannot be found.
%
\begin{code}
  let rec assoc (d, k, l) = match l with
    | (ki, vi)::tl ->
       if ki = k then
         vi
       else
         assoc (d, k, tl)
    | _ -> d

@@

>>>>

The remaining 473 programs for which we could not produce a witness did
not admit such an automatic diagnosis, so we selected a random sample of
50 programs and manually searched for a witness.
%
We categorized the programs into four groups.

\paragraph{Witness Exists}
%
We found only eight cases (or 16\% of the samples) where a witness existed,
but \toolname was unable to discover it.
%
Five involved synthesizing a \emph{pair} of specially-crafted inputs
that would result in the function returning values of incompatible
types.
%
The other three required synthesizing an input that would trigger a
particular path through the program, and would likely have been caught
by symbolic execution.
%
% Overall, it appears that our random instantiation of holes is well-suited
% to finding type errors in student programs.
%
%We categorized the remaining 42 failures into three groups.

@@

\paragraph{Bad Function Calls}
Fourteen programs (or 28\% of the samples) contained a function call
where the student had supplied incorrect inputs, yet no error was
triggered.
%
Consider, for example the following |assoc| function, which
looks up a key in an \emph{association list} and returns a default if
the key cannot be found.
%
\begin{code}
  let rec assoc (d, k, l) = match l with
    | (ki, vi)::tl ->
       if ki = k then
         vi
       else
         assoc (d, k, tl)
    | _ -> d

@@

To summarize, our investigation suggests that the majority of programs
(88\%) for which we fail to find a witness do not, in fact, admit a
witness.
%
Overall, it appears that our random instantiation of holes is well-suited
to finding type errors in student programs.
%
The remaining programs were all cases where \ocaml's type system was
overly conservative.
%
Of course, the conservatism is somewhat justified as each case pointed
to a ``code smell'' or difficult-to-use API; it would be interesting to
investigate how demonstrate these issues in a similar manner.
